Practical Considerations for Lomb-Scargle Periodic Analysis

### 2015-08-24 paper.tex
more filling-in of outline

\todo{Quick motivating example with code \& figures}

@@

@@

A particularly special function in the complex exponential, $e^{2\pi i\nu_0 t}$, which has the Fourier transform $\mathcal{F}\{e^{2\pi i\nu_0 t}\} = \delta(\nu - \nu_0)$, which follows directly from the definition of the Dirac delta function:

@@

This is the sense in which a Fourier power is an ideal measure of periodicity: the Fourier transform of a pure periodic signal is a single localized spike identifying its frequency.

@@

\todo{talk about offsets and phase-shifts?}

With this in mind, we can define the Fourier power,

\begin{equation}
  P(\nu) = \left|\hat{g}(\nu)\right|^2,
\end{equation}

which measures the (real) squared-magnitude of the (complex) Fourier transform.

\todo{Some examples of functions, their transforms, and their power.}

\todo{convolution function and windowing}

@@

\todo{Evenly-spaced delta functions $\to$ evenly-spaced delta in frequency, separated by nyquist}

@@

\todo{Unevenly-spaced delta functions: no cancelations, and so window is generally {\it infinitely} wide.}

\todo{Least-squares equivalence of Lomb-Scargle}

@@

\todo{Follow discussion from gatspy documentation}

@@

\todo{Talk about peak width and False Alarm Probability; recommend bootstrap (show example of when analytic FAP fails?)}

@@

\todo{Talk about Press \& Rybicki method; show some benchmarks.}

@@

\todo{Multiterm makes the true period fit better, but also bumps the background noise.}

>>>>

[Quick motivating example with figures: RR Lyrae star from LINEAR?]

@@

[Perhaps put this info in an appendix?]

@@

\subsubsection{Fourier Transform and Periodic Functions}

A particularly special function in the complex exponential, $e^{2\pi i\nu_0 t} \equiv \cos(2\pi \nu_0 t) + i\sin(2\pi\nu_0 t)$, which has the Fourier transform $\mathcal{F}\{e^{2\pi i\nu_0 t}\} = \delta(\nu - \nu_0)$, which follows directly from the definition of the Dirac delta function:

@@

Similarly, if we have a signal composed of a sum of complex exponentials, the linearity of the Fourier transform assures that the transform is simply a sum of delta functions:
\begin{equation}
  \mathcal{F}\left\{ \sum_{n=1}^N e^{-2\pi i \nu_n t} \right\} = \sum_{n=1}^N \delta(\nu - \nu_n)
\end{equation}
This result shows the sense in which a Fourier transform is an ideal measure of periodicity: the Fourier transform pulls-out and identifies the individual periodic components of any input.

@@

Using the above definitions, it is straightforward to show that a time-offset in a function leads to a multiplicative phase in the Fourier transform; that is, if
\begin{equation}
  g_\tau(t) \equiv g(t - \tau)
\end{equation}
then 
\begin{equation}
  \mathcal{F}\{g_\tau\} = e^{-2\pi i \nu \tau}\mathcal{F}\{g\}
\end{equation}
Because the phase merely represents the chosen coordinate system, we'd like to remove it for spectral analysis.
One way to do this is to compute the power spectral density, which is proportional to \todo{const of proportionality?} the squared modulus of the Fourier Transform:
\begin{equation}
  P\{g\} \propto \left| \mathcal{F}\{g\} \right|^2.
\end{equation}
Because the modulus of a phase is equal to unity, the resulting measure is unchanged under time-translations in $g(t)$.

\begin{itemize}
\item Some examples of functions, their transforms, and their power.
\end{itemize}

\subsection{Effect of Window Functions}

\begin{itemize}
\item convolution function and windowing
\end{itemize}

@@

\begin{itemize}
\item Evenly-spaced delta functions $\to$ evenly-spaced delta in frequency, separated by nyquist
\end{itemize}

@@

\begin{itemize}
\item Unevenly-spaced delta functions: no cancelations, and so window is generally {\it infinitely} wide.
\item Least-squares equivalence of Lomb-Scargle
\end{itemize}

@@

\begin{itemize}
\item Follow discussion from gatspy documentation
\end{itemize}

@@

\begin{itemize}
\item Talk about peak width and False Alarm Probability; recommend bootstrap (show example of when analytic FAP fails?)
\end{itemize}

@@

\begin{itemize}
\item Talk about Press \& Rybicki method; show some benchmarks.
\end{itemize}

@@

\begin{itemize}
\item Multiterm makes the true period fit better, but also bumps the background noise.
\end{itemize}
